# Repdrocuing
In this part we will be reproducing an analysis with data supplied by J. Louter (INT/ILC)
Different compounds were tested on adult C.elegans and the amount of offspring were counted.

## Getting the data
First we will need to get the data from somewhere, we use wget to download the file excel file from github.
```{bash include=FALSE}
# Go to data_raw
cd data_raw/ 

# Download the excel sheet dataset
wget -O "CE.LIQ.FLOW.062_Tidydata.xlsx" https://github.com/DataScienceILC/tlsc-dsfb26v-20_workflows/raw/main/data/CE.LIQ.FLOW.062_Tidydata.xlsx

```
Now we can read the file into R.
```{r}
# Load libraries
library(readxl)
library(dplyr)

# Read the excel file
C.elegans_data <- read_excel(path = "data_raw/CE.LIQ.FLOW.062_Tidydata.xlsx")

# Display the important rows
C.elegans_data %>% select(c(RawData, compName, compConcentration))

```
Here we see that compConcentration is of chr type, so we have to change it into a numeric type.
```{r}
# Load readr
library(readr)

# Use parse_number to change the conecentration into numeric
C.elegans_data$compConcentration <- parse_number(C.elegans_data$compConcentration)

# Display the rows again
C.elegans_data %>% select(c(RawData, compName, compConcentration))
```

```{r}
# Load required libraries
library(ggplot2)

# Create scatterplot
ggplot(C.elegans_data, aes(x = compConcentration, y = RawData, color = compName, shape = expType)) +
  geom_point(position = position_jitter(width = 0.2, height = 0)) +
  labs(x = "Compound Concentration", y = "Raw Data Counts") +
  scale_x_continuous(labels = scales::comma_format()) +  # Fix x-axis labels for readability
  theme_minimal()  # Optional: Customize the theme if needed

```
The graph is really cramped into the left part because of a couple of very high concentrations. This occurs because the concentrations are created by diluting the previous solution by 1/10.


